---
format:
  docx:
    number-sections: true
    toc: true
execute:
  echo: false
  message: false
  warning: false
engine: knitr
fig-dpi: 400
---

# Setup

```{r setup}
setwd("D:/Clients/VinUni/Dr. Linh/Dietary_index")
library(dplyr)
library(DataExplorer)
library(knitr)
library(tidyr)
library(ggplot2)
library(forcats)
library(haven)
library(foreign)
library(readr)
library(purrr)
library(dietaryindex)
library(writexl)
library(nhanesA)
library(naniar)
library(VIM)
library(gridExtra)
library(scales)
library(gt)
library(tidyverse)
library(naniar)    # missing summaries 
library(mice)      # md.pattern, missing clustering
library(janitor)   
library(UpSetR)    # upset plot cho pattern
library(visdat)    
library(skimr)     # skim for quick summary
library(flextable)
library(splines)
library(labelled)
library(gtsummary)
library(EValue)
library(survey)

# Clear existing data and graphics
rm(list=ls())
graphics.off()

# Define cycles and their NHANES letter suffixes 
cycles  <- c("0506","0708","0910","1112","1314","1516","1718")
letters <- c("D","E","F","G","H","I","J")
```

# Functions

```{r functions}
# A function to average the day 1 and day 2 data
NHANES_d1d2_average = function(COHORT1, COHORT2, identifier){
    # Check which columns exist in both COHORT and COHORT2
    common_cols <- intersect(colnames(COHORT1), colnames(COHORT2))

    # Remove 'SEQN' as it is the joining key, not an average-able variable
    common_cols <- setdiff(common_cols, identifier)

    # Dynamically generate the mutate expressions for averaging the common columns between the two data sets
    avg_exprs <- setNames(lapply(common_cols, function(col) {
        rlang::parse_expr(paste0(col, " = (", col, ".x + ", col, ".y) / 2"))
    }), common_cols)

    # Perform inner join to merge the two data sets
    COHORT12 <- inner_join(COHORT1, COHORT2, by = identifier)

    # Perform the averaging calculations
    COHORT12 <- COHORT12 %>%
        mutate(!!!avg_exprs)

    # Explicitly select the columns of interest
    COHORT12 <- COHORT12 %>%
        dplyr::select(identifier, !!!common_cols)
    
    # Return the averaged data set
    return(COHORT12)
}

# A funtion for tab table (n, %, included NA)
gt_cov_2col <- function(data, vars, digits = 1, title = NULL) {
  stopifnot(is.data.frame(data), is.character(vars))
  
  tab <- map_dfr(vars, function(v) {
    x <- data[[v]]
    if (is.null(x)) stop(sprintf("Variable '%s' not found in data.", v))
    n_total <- length(x)
    
    df <- tibble(level_chr = as.character(x), is_na = is.na(x)) %>%
      mutate(level_chr = ifelse(is_na, "NA", level_chr)) %>%
      count(level_chr, is_na, name = "n") %>%
      arrange(is_na, desc(n)) %>%
      mutate(
        variable = v,
        percent  = 100 * n / n_total
      ) %>%
      transmute(variable, level = level_chr, n, percent)
  })
  
  tab %>%
    gt(groupname_col = "variable") %>%
    cols_label(
      level   = "Level",
      n       = md("**n**"),
      percent = md("**%**")
    ) %>%
    fmt_number(columns = percent, decimals = digits) %>%
    tab_options(table.font.size = px(12), data_row.padding = px(4))
}

# Function: download XPT and keep SEQN + specified columns
read_keep <- function(url, keep) {
  tf <- tempfile(fileext = ".XPT")
  download.file(url, tf, mode = "wb", quiet = TRUE)
  df <- foreign::read.xport(tf)
  df[, intersect(names(df), c("SEQN", keep)), drop = FALSE]
}

# Function: left join while dropping duplicate columns from y
left_join_drop_dups <- function(x, y, by) {
  cols_y <- union(by, setdiff(names(y), names(x)))
  y_sel  <- select(y, all_of(cols_y))
  left_join(x, y_sel, by = by)
}

FitFlextableToPage <- function(ft, pgwidth = 6){
  ft_out <- ft |> autofit()  # Automatically adjust columns to content width
  ft_out <- width(ft_out, width = dim(ft_out)$widths * pgwidth / (flextable_dim(ft_out)$widths))
  return(ft_out)
}

fisher.test.simulate.p.values <- function(data, variable, by, ...) {
  result <- list()
  test_results <- stats::fisher.test(data[[variable]], data[[by]], simulate.p.value = TRUE)
  result$p <- test_results$p.value
  result$test <- test_results$method
  result
}
```

# Data preprocessing

## Demo & health variables

```{r download and merge data}
#| eval: false
#| include: false

years   <- c("2005","2007","2009","2011","2013","2015","2017")
suffix  <- c("_D","_E","_F","_G","_H","_I","_J")
cycles  <- c("0506","0708","0910","1112","1314","1516","1718")

# Groups of variables
demo_cols   <- c("RIAGENDR","RIDAGEYR","SDMVSTRA","SDMVPSU","WTMEC2YR", "WTDR2D", "DMDEDUC2","RIDRETH1","DMDMARTL","INDFMPIR", "DMDHHSIZ")
mcq_cols    <- c("MCQ160B","MCQ160C","MCQ160D","MCQ160E","MCQ160F","MCQ220")
diq_cols    <- c("DIQ010","DIQ050", "DID060", "DIQ070", "DID040", "DIQ175S")
bpq_cols    <- c("BPQ020")
bpx_cols    <- c("BPXSY1","BPXSY2","BPXSY3","BPXSY4","BPXDI1","BPXDI2","BPXDI3","BPXDI4") # full cycles have four readings
bmx_cols    <- c("BMXBMI","BMXWAIST")
ghb_cols    <- c("LBXGH")
glu_cols    <- c("LBXGLU","LBXIN")     # LBXIN: Insulin (uU/mL)
trigly_cols <- c("LBDLDL","LBXTR","LBDAPB")
hdl_cols    <- c("LBDHDD")

# covariate components
smq_cols    <- c("SMQ020","SMQ040")      # smoking 
alq_cols    <- c("ALQ130")               # alcohol drinks/day (12 mo)
paq_cols    <- c("PAQ670","PAD675")      # days/wk * minutes/day 
#dsd_cols    <- c("DSD010")               # vitamin/mineral supplement use
rhq_cols    <- c("RHQ162")               # gestational diabetes

# Download and merge all health-related files for each cycle
per_cycle <- vector("list", length(cycles))
names(per_cycle) <- cycles

for (i in seq_along(years)) {
  yr  <- years[i]; suf <- suffix[i]
  
  # Short function to build NHANES file URL
  U <- function(stub) sprintf("https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/%s/DataFiles/%s%s.XPT", yr, stub, suf)
  
  # Read each component
  demo_i <- read_keep(U("DEMO"),   demo_cols)
  mcq_i  <- read_keep(U("MCQ"),    mcq_cols)
  diq_i  <- read_keep(U("DIQ"),    diq_cols)
  bpq_i  <- read_keep(U("BPQ"),    bpq_cols)
  bpx_i  <- read_keep(U("BPX"),    bpx_cols)
  bmx_i  <- read_keep(U("BMX"),    bmx_cols)
  ghb_i  <- read_keep(U("GHB"),    ghb_cols)
  glu_i  <- read_keep(U("GLU"),    glu_cols)
  trg_i  <- read_keep(U("TRIGLY"), trigly_cols)
  hdl_i  <- read_keep(U("HDL"),    hdl_cols)
  
  smq_i  <- read_keep(U("SMQ"),    smq_cols)
  alq_i  <- read_keep(U("ALQ"),    alq_cols)
  paq_i  <- read_keep(U("PAQ"),    paq_cols)
  rhq_i  <- read_keep(U("RHQ"),    rhq_cols)
  
  # chưa merge dsd_cols: check lại link path
  
  # Merge into one dataframe for this cycle
  per_cycle[[i]] <- Reduce(
    function(x, y) merge(x, y, by = "SEQN", all.x = TRUE, sort = FALSE),
    list(demo_i, mcq_i, diq_i, bpq_i, bpx_i, bmx_i, ghb_i, glu_i, trg_i, hdl_i, smq_i, alq_i, paq_i, rhq_i)
  )
}

# Bind all cycles and add "cycle" column
demo_health_0518 <- bind_rows(per_cycle, .id = "cycle")
glimpse(demo_health_0518)
# View(demo_health_0518)

# supplement "vitamin/mineral supplement use" (DSD010) variable
dsd_table <- c("DSQ1_D", "DSQTOT_E", "DSQTOT_F", "DSQTOT_G", "DSQTOT_H", "DSQTOT_I", "DSQTOT_J")
  
dsd_list <- list ()
for (tb in dsd_table) {
  df <- nhanes(
    nh_table = tb,
    includelabels = FALSE,
    translated = TRUE,
    cleanse_numeric = FALSE,
    nchar = 128,
    adjust_timeout = TRUE
    )
  dsd_list[[tb]] <- df
}

dsd_0518 <- bind_rows(dsd_list)
dsd_0518 <- select(dsd_0518, c("SEQN", "DSD010"))

# merge dsd in "demo_health_0518" df
demo_health_0518 <- demo_health_0518 %>%
  left_join(dsd_0518, by = "SEQN")

# Save demo_health_0518 to CSV file
write_csv(demo_health_0518, "DATA/demo_health_0518.csv")

# Clear from memory and load back from CSV
rm(demo_health_0518, dsd_0518, dsd_list)
gc()
```

```{r load demo_health_0518}
# Load demo_health_0518 back from CSV
demo_health_0518 <- read_csv("D:/Clients/VinUni/Dr. Linh/Dietary_index/DATA/demo_health_0518.csv", show_col_types = FALSE)
```

## Dietary index variables

```{r prepare common data}
#| eval: false
#| include: false
# Create an empty list to store all cycles
NHANES_ALL <- list()

# Loop: read files for each cycle and store as a nested list
for (i in seq_along(cycles)) {
  cyc <- cycles[i]
  let <- letters[i]
  
  # Build subfolders for this cycle
  fped_dir <- file.path("DATA", "FPED",     paste0("FPED_", cyc))
  nutr_dir <- file.path("DATA", "NUTRIENT", paste0("NUTRIENT_", cyc))
  demo_dir <- file.path("DATA", "DEMO",     paste0("DEMO_", cyc))
  
  # Read 9 expected tables for the cycle
  # FPED: day 1/2 total; NUTRIENT: day 1/2 total; NUTRIENT_IND/FPED_IND: individual foods
  cycle_list <- list(
    DEMO          = read_xpt(file.path(demo_dir, sprintf("DEMO_%s.xpt", let))),             
    
    FPED          = read_sas(file.path(fped_dir, sprintf("fped_dr1tot_%s.sas7bdat", cyc))), 
    NUTRIENT      = read_xpt(file.path(nutr_dir, sprintf("DR1TOT_%s.xpt", let))),          
    
    FPED2         = read_sas(file.path(fped_dir, sprintf("fped_dr2tot_%s.sas7bdat", cyc))), 
    NUTRIENT2     = read_xpt(file.path(nutr_dir, sprintf("DR2TOT_%s.xpt", let))),          
    
    NUTRIENT_IND  = read_xpt(file.path(nutr_dir, sprintf("DR1IFF_%s.xpt", let))),           
    FPED_IND      = read_sas(file.path(fped_dir, sprintf("fped_dr1iff_%s.sas7bdat", cyc))), 
    
    FPED_IND2     = read_sas(file.path(fped_dir, sprintf("fped_dr2iff_%s.sas7bdat", cyc))), 
    NUTRIENT_IND2 = read_xpt(file.path(nutr_dir, sprintf("DR2IFF_%s.xpt", let)))            
  )
  
  # Save this cycle into the outer list, named by its code (e.g., "1718")
  NHANES_ALL[[cyc]] <- cycle_list
}
```

### HEI2020

```{r prepare and calculate HEI2020}
#| eval: false
#| include: false
# Prepare HEI2020 specific data WITH weight variable
HEI2020_0518 <- map_dfr(cycles, function(cyc) { 
  dat <- NHANES_ALL[[cyc]]
  
  # Calculate HEI scores
  out <- HEI2020_NHANES_FPED(
    FPED_PATH      = dat$FPED,
    NUTRIENT_PATH  = dat$NUTRIENT,
    DEMO_PATH      = dat$DEMO,
    FPED_PATH2     = dat$FPED2,
    NUTRIENT_PATH2 = dat$NUTRIENT2
  )
  
  # Extract weight variable from NUTRIENT data
  weights <- dat$NUTRIENT %>% 
    select(SEQN, WTDR2D)
  
  # Merge weight and add cycle
  out %>%
    left_join(weights, by = "SEQN") %>%
    mutate(cycle = cyc)
})

#View(HEI2020_0518)
```

### AHEI

```{r prepare and calculate AHEI}
#| eval: false
#| include: false
# Prepare AHEI specific data
AHEI_0518 <- map_dfr(cycles, function(cyc) {
  dat <- NHANES_ALL[[cyc]]

  # Calculate AHEI scores
  out <- AHEI_NHANES_FPED(
    FPED_IND_PATH      = dat$FPED_IND,
    NUTRIENT_IND_PATH  = dat$NUTRIENT_IND,
    FPED_IND_PATH2     = dat$FPED_IND2,
    NUTRIENT_IND_PATH2 = dat$NUTRIENT_IND2
  )
  
  # Extract weight variable from NUTRIENT data
  weights <- dat$NUTRIENT %>% 
    select(SEQN, WTDR2D)
  
  # Merge weight and add cycle
  out %>%
    left_join(weights, by = "SEQN") %>%
    mutate(cycle = cyc)
})

#View(AHEI_0518)
```

### DASH

```{r prepare and calculate DASH}
#| eval: false
#| include: false
# Prepare DASH specific data
DASH_0518 <- map_dfr(cycles, function(cyc) {
  dat <- NHANES_ALL[[cyc]]

  # Calculate DASH scores
  out <- DASH_NHANES_FPED(
    FPED_IND_PATH      = dat$FPED_IND,
    NUTRIENT_IND_PATH  = dat$NUTRIENT_IND,
    FPED_IND_PATH2     = dat$FPED_IND2,
    NUTRIENT_IND_PATH2 = dat$NUTRIENT_IND2
  )

  # Extract weight variable from NUTRIENT data
  weights <- dat$NUTRIENT %>% 
    select(SEQN, WTDR2D)
  
  # Merge weight and add cycle
  out %>%
    left_join(weights, by = "SEQN") %>%
    mutate(cycle = cyc)
})

#View(DASH_0518)
```

### aMED

```{r prepare and calculate aMED}
#| eval: false
#| include: false
# Prepare aMED specific data
AMED_0518 <- map_dfr(cycles, function(cyc) {
  dat <- NHANES_ALL[[cyc]]

  # Calculate aMED scores
  out <- MED_NHANES_FPED(
    FPED_PATH      = dat$FPED,
    NUTRIENT_PATH  = dat$NUTRIENT,
    DEMO_PATH      = dat$DEMO,
    FPED_PATH2     = dat$FPED2,
    NUTRIENT_PATH2 = dat$NUTRIENT2
  )

  # Extract weight variable from NUTRIENT data
  weights <- dat$NUTRIENT %>% 
    select(SEQN, WTDR2D)
  
  # Merge weight and add cycle
  out %>%
    left_join(weights, by = "SEQN") %>%
    mutate(cycle = cyc)
})

#View(AMED_0518)
```

### DII

```{r prepare and calculate DII}
#| eval: false
#| include: false
# Prepare DII specific data
DII_0518 <- map_dfr(cycles, function(cyc) {
  dat <- NHANES_ALL[[cyc]]

  # Calculate DII scores
  out <- DII_NHANES_FPED(
    FPED_PATH      = dat$FPED,
    NUTRIENT_PATH  = dat$NUTRIENT,
    DEMO_PATH      = dat$DEMO,
    FPED_PATH2     = dat$FPED2,
    NUTRIENT_PATH2 = dat$NUTRIENT2
  )

  # Extract weight variable from NUTRIENT data
  weights <- dat$NUTRIENT %>% 
    select(SEQN, WTDR2D)
  
  # Merge weight and add cycle
  out %>%
    left_join(weights, by = "SEQN") %>%
    mutate(cycle = cyc)
})

#View(DII_0518)
```

### PHDI

```{r prepare and calculate PHDI}
#| eval: false
#| include: false
# ---------- PHDI ----------
# prepare data (PHDI function requires SERV_DATA=FPED_NUTRIENT_DEMO in a dataset) 
# DR1TKCAL: Energy (kcal) - Dietary Interview - Total Nutrient Intakes, First Day
# Dietary recall status: DR1DRSTZ == 1 OR DR2DRSTZ == 1 (Reliable and met the minimum criteria) 

# --------- link the food pattern data for each NHANES individual with their nutrient and demographic data ---------
# import food pattern (unit grams/day)

# Read in the data
# FPED file (food pattern variable files)
## 2017-2018
FPED_DR1TOT_gram_1718_d1 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_1718_d1.csv")
FPED_DR1TOT_gram_1718_d2 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_1718_d2.csv")
## 2015-2016
FPED_DR1TOT_gram_1516_d1 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_1516_d1.csv")
FPED_DR1TOT_gram_1516_d2 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_1516_d2.csv")
## 2013-2014
FPED_DR1TOT_gram_1314_d1 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_1314_d1.csv")
FPED_DR1TOT_gram_1314_d2 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_1314_d2.csv")
## 2011-2012
FPED_DR1TOT_gram_1112_d1 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_1112_d1.csv")
FPED_DR1TOT_gram_1112_d2 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_1112_d2.csv")
## 2009-2010
FPED_DR1TOT_gram_0910_d1 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_0910_d1.csv")
FPED_DR1TOT_gram_0910_d2 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_0910_d2.csv")
## 2007-2008
FPED_DR1TOT_gram_0708_d1 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_0708_d1.csv")
FPED_DR1TOT_gram_0708_d2 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_0708_d2.csv")
## 2005-2006
FPED_DR1TOT_gram_0506_d1 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_0506_d1.csv")
FPED_DR1TOT_gram_0506_d2 = read_csv("DATA/FPED_DR1TOT_gram/FPED_DR1TOT_gram_0506_d2.csv")

NUTRIENT_DEMO_list <- list ()
for (cyc in cycles) {
  df <- NHANES_ALL[[cyc]]
  df_merged <- reduce(                            # merge the tables in the list one by one
    list(
    df$DEMO,
    df$NUTRIENT %>% select(SEQN, DR1TKCAL, DR1DRSTZ, WTDR2D),   #doulbe-check DR1DRSTZ/DR2DRSTZ
    df$NUTRIENT2 %>% select(SEQN, DR2TKCAL, DR2DRSTZ)),
  ~ left_join(.x, .y, by = "SEQN"))
  NUTRIENT_DEMO_list[[cyc]] <- df_merged
}

#FPED_NUTRIENT_DEMO_list_d1 <- left_join (NUTRIENT_DEMO_list[[0506]], FPED_DR1TOT_gram_0506_d1 , by = "SEQN")
#FPED_NUTRIENT_DEMO_list_d2 <- left_join (NUTRIENT_DEMO_list[[0506]], FPED_DR1TOT_gram_0506_d2 , by = "SEQN")

# Create list to save data
FPED_NUTRIENT_DEMO_list_d1 <- list()
FPED_NUTRIENT_DEMO_list_d2 <- list()

# loop for merging data
for (cyc in cycles) {

  # lấy dữ liệu DEMO+NUTRIENT từ list
  demo_data <- NUTRIENT_DEMO_list[[cyc]]

  # lấy dữ liệu FPED tương ứng theo tên biến
  fped_d1 <- get(paste0("FPED_DR1TOT_gram_", cyc, "_d1"))
  fped_d2 <- get(paste0("FPED_DR1TOT_gram_", cyc, "_d2"))

  # ghép: giữ toàn bộ hàng trong DEMO
  FPED_NUTRIENT_DEMO_list_d1[[cyc]] <- left_join(demo_data, fped_d1, by = "SEQN")
  FPED_NUTRIENT_DEMO_list_d2[[cyc]] <- left_join(demo_data, fped_d2, by = "SEQN")
}

# tibble 
FPED_NUTRIENT_DEMO_list_d1[["1516"]] %>% glimpse()

#FPED_NUTRIENT_DEMO_0518_d1 <- bind_rows(FPED_NUTRIENT_DEMO_d1_list)

# ------------------ Calculate PHDI ------------------------------------------
# PHDI day 1
PHDI_0518_d1 <- map_dfr(cycles, function(cyc) {
  dat <- FPED_NUTRIENT_DEMO_list_d1[[cyc]]
  out <- PHDI_V2(
    SERV_DATA = dat,
    RESPONDENTID = dat$SEQN,
    GENDER = dat$RIAGENDR, # check RIAGENDR 
    TOTALKCAL_PHDI = dat$DR1TKCAL,
    WGRAIN_SERV_PHDI = dat$F_WHOLE_gram, 
    STARCHY_VEG_SERV_PHDI = dat$V_STARCHY_TOTAL_gram, 
    VEG_SERV_PHDI = dat$V_TOTAL_NONSTARCHY_gram, 
    FRT_SERV_PHDI = dat$F_WHOLE_gram, 
    DAIRY_SERV_PHDI = dat$D_TOTAL_gram, 
    REDPROC_MEAT_SERV_PHDI = dat$PF_REDMEAT_gram, 
    POULTRY_SERV_PHDI = dat$PF_POULT_gram, 
    EGG_SERV_PHDI = dat$PF_EGGS_gram, 
    FISH_SERV_PHDI = dat$PF_SEAFD_TOTAL_gram, 
    NUTS_SERV_PHDI = dat$PF_NUTSDS_gram, 
    LEGUMES_SERV_PHDI = dat$PF_LEGUMES_gram, 
    SOY_SERV_PHDI = dat$PF_SOY_gram, 
    ADDED_FAT_UNSAT_SERV_PHDI = dat$OILS_gram, 
    ADDED_FAT_SAT_TRANS_SERV_PHDI = dat$SOLID_FATS, 
    ADDED_SUGAR_SERV_PHDI = dat$ADD_SUGARS_gram
  )

  # Extract weight variable from NUTRIENT data
  weights <- dat %>%
    select(SEQN, WTDR2D) %>%
    rename(RESPONDENTID = SEQN)
  
  # Merge weight and add cycle
  out %>%
    left_join(weights, by = "RESPONDENTID") %>%
    mutate(cycle = cyc)
})

# PHDI day 2
PHDI_0518_d2 <- map_dfr(cycles, function(cyc) {
  dat <- FPED_NUTRIENT_DEMO_list_d2[[cyc]]
  out <- PHDI_V2(
    SERV_DATA = dat,
    RESPONDENTID = dat$SEQN,
    GENDER = dat$RIAGENDR,
    TOTALKCAL_PHDI = dat$DR2TKCAL,
    WGRAIN_SERV_PHDI = dat$F_WHOLE_gram, 
    STARCHY_VEG_SERV_PHDI = dat$V_STARCHY_TOTAL_gram, 
    VEG_SERV_PHDI = dat$V_TOTAL_NONSTARCHY_gram, 
    FRT_SERV_PHDI = dat$F_WHOLE_gram, 
    DAIRY_SERV_PHDI = dat$D_TOTAL_gram, 
    REDPROC_MEAT_SERV_PHDI = dat$PF_REDMEAT_gram, 
    POULTRY_SERV_PHDI = dat$PF_POULT_gram, 
    EGG_SERV_PHDI = dat$PF_EGGS_gram, 
    FISH_SERV_PHDI = dat$PF_SEAFD_TOTAL_gram, 
    NUTS_SERV_PHDI = dat$PF_NUTSDS_gram, 
    LEGUMES_SERV_PHDI = dat$PF_LEGUMES_gram, 
    SOY_SERV_PHDI = dat$PF_SOY_gram, 
    ADDED_FAT_UNSAT_SERV_PHDI = dat$OILS_gram, 
    ADDED_FAT_SAT_TRANS_SERV_PHDI = dat$SOLID_FATS, 
    ADDED_SUGAR_SERV_PHDI = dat$ADD_SUGARS_gram
  )

  # Extract weight variable from NUTRIENT data
  weights <- dat %>%
    select(SEQN, WTDR2D) %>%
    rename(RESPONDENTID = SEQN)
  
  # Merge weight and add cycle
  out %>%
    left_join(weights, by = "RESPONDENTID") %>%
    mutate(cycle = cyc)
})

# Average PHDI (PHDI_0518 <- (PHDI_0518_d1 + PHDI_0518_d2)/2)
# merge PHDI day 1 + day 2
PHDI_0518_d1_d2 <- full_join(
  PHDI_0518_d1,
  PHDI_0518_d2,
  by = c("RESPONDENTID", "GENDER", "cycle", "WTDR2D"),
  suffix = c("_d1", "_d2")
)

# List of 'PHDI_' vars (filter list of vars started as "PHDI...") 
phdi_vars <- grep("^PHDI_", names(PHDI_0518_d1), value = TRUE) 
phdi_vars <- c(phdi_vars, "TOTALKCAL_PHDI")

# Loop to compute mean each var PHDI_
for (v in phdi_vars) {
  v_d1 <- paste0(v, "_d1")
  v_d2 <- paste0(v, "_d2")
  v_mean <- paste0(v, "_mean")
  
  PHDI_0518_d1_d2[[v_mean]] <- rowMeans(
    cbind(PHDI_0518_d1_d2[[v_d1]], PHDI_0518_d1_d2[[v_d2]]),
    na.rm = TRUE
  )
}

glimpse(PHDI_0518_d1_d2)

PHDI_0518_mean <- PHDI_0518_d1_d2 %>%
  select(RESPONDENTID, WTDR2D, GENDER, cycle, ends_with("_mean")) %>%
  rename(SEQN = RESPONDENTID)

glimpse(PHDI_0518_mean)
```

## Mortality variables

```{r}
#| eval: false
#| include: false
# Map cycle 
cyc_years <- c(
  "0506" = "2005_2006",
  "0708" = "2007_2008",
  "0910" = "2009_2010",
  "1112" = "2011_2012",
  "1314" = "2013_2014",
  "1516" = "2015_2016",
  "1718" = "2017_2018"
)

# Function to read mortality data
read_mort <- function(dir, cyc) {
  yrs  <- cyc_years[[cyc]]
  path <- file.path(dir, paste0("NHANES_", yrs, "_MORT_2019_PUBLIC.dat"))
  out <- read_fwf(
    file      = path,
    col_types = "iiiiiiii",
    col_positions = fwf_cols(
      SEQN         = c(1, 6),
      eligstat     = c(15, 15),
      mortstat     = c(16, 16),
      ucod_leading = c(17, 19),
      diabetes     = c(20, 20),
      hyperten     = c(21, 21),
      permth_int   = c(43, 45),
      permth_exm   = c(46, 48)
    ),
    na = c("", ".")
  )
  mutate(out, cycle = cyc)
}

# Read and merge mortality data for 2005-2018
MORT_DIR <- "DATA/Mortality"
cycles   <- names(cyc_years)

mort_0518 <- do.call(
  bind_rows,
  lapply(cycles, function(cyc) read_mort(MORT_DIR, cyc))
)
#View(mort_0518)
```

## Greenhouse gas variable

```{r}
#| eval: false
#| include: false
# Use only the dietary recall that is reliable and met the minimum criteria in NHANES-DR1DRSTZ ==1 OR DR2DRSTZ == 1 (James applied)
# import GHG day 1 and day 2 data
# Set base_path folder 
BASE <- "DATA/GHG"

# create a blank frame
GHG_0518 <- data.frame()

# loop: import data + average day1-day2
for (cyc in cycles) {
  # build file paths
  d1_path <- file.path (BASE, paste0("GHG_DR1TOT_gram_", cyc, ".csv"))
  d2_path <- file.path (BASE, paste0("GHG_DR2TOT_gram_", cyc, ".csv"))
  
  # import data
  d1 <- read_csv(d1_path)
  d2 <- read_csv(d2_path)
  
  # Average day1–day2
  avg <- NHANES_d1d2_average(d1, d2, identifier = "SEQN")
  
  # Add cycle label
  avg <- avg %>% mutate(cycle = cyc)
  
  # combined dataset
  GHG_0518 <- bind_rows(GHG_0518, avg)
}

glimpse(GHG_0518)
# View(GHG_0518)

#GHG_DR1TOT_gram_1718_d1 = read_csv("GHG_DR1TOT_gram_1718.csv")
#GHG_DR1TOT_gram_1718_d2 = read_csv("GHG_DR2TOT_gram_1718.csv")
#GHG_DR1TOT_gram_1516_d1 = read_csv("GHG_DR1TOT_gram_1516.csv")
#GHG_DR1TOT_gram_1516_d2 = read_csv("GHG_DR2TOT_gram_1516.csv")
#GHG_DR1TOT_gram_1314_d1 = read_csv("GHG_DR1TOT_gram_1314.csv")
#GHG_DR1TOT_gram_1314_d2 = read_csv("GHG_DR2TOT_gram_1314.csv")
#GHG_DR1TOT_gram_1112_d1 = read_csv("GHG_DR1TOT_gram_1112.csv")
#GHG_DR1TOT_gram_1112_d2 = read_csv("GHG_DR2TOT_gram_1112.csv")
#GHG_DR1TOT_gram_0910_d1 = read_csv("GHG_DR1TOT_gram_0910.csv")
#GHG_DR1TOT_gram_0910_d2 = read_csv("GHG_DR2TOT_gram_0910.csv")
#GHG_DR1TOT_gram_0708_d1 = read_csv("GHG_DR1TOT_gram_0708.csv")
#GHG_DR1TOT_gram_0708_d2 = read_csv("GHG_DR2TOT_gram_0708.csv")
#GHG_DR1TOT_gram_0506_d1 = read_csv("GHG_DR1TOT_gram_0506.csv")
#GHG_DR1TOT_gram_0506_d2 = read_csv("GHG_DR2TOT_gram_0506.csv")

# Merge the day 1 and day 2 data for each cohort
#GHG_1718_d1d2 = NHANES_d1d2_average(GHG_DR1TOT_gram_1718_d1, GHG_DR1TOT_gram_1718_d2, identifier = "SEQN")
#GHG_1516_d1d2 = NHANES_d1d2_average(GHG_DR1TOT_gram_1516_d1, GHG_DR1TOT_gram_1516_d2, identifier = "SEQN")
#GHG_1314_d1d2 = NHANES_d1d2_average(GHG_DR1TOT_gram_1314_d1, GHG_DR1TOT_gram_1314_d2, identifier = "SEQN")
#GHG_1112_d1d2 = NHANES_d1d2_average(GHG_DR1TOT_gram_1112_d1, GHG_DR1TOT_gram_1112_d2, identifier = "SEQN")
#GHG_0910_d1d2 = NHANES_d1d2_average(GHG_DR1TOT_gram_0910_d1, GHG_DR1TOT_gram_0910_d2, identifier = "SEQN")
#GHG_0708_d1d2 = NHANES_d1d2_average(GHG_DR1TOT_gram_0708_d1, GHG_DR1TOT_gram_0708_d2, identifier = "SEQN")
#GHG_0506_d1d2 = NHANES_d1d2_average(GHG_DR1TOT_gram_0506_d1, GHG_DR1TOT_gram_0506_d2, identifier = "SEQN")
```

## Merging all data

```{r}
#| eval: false
#| include: false
by_keys <- c("SEQN", "cycle")

# HEALTH -> (HEI, AHEI, DASH, aMED, DII, PHDI_mean) -> MORT -> GHG_0518
nhanes_merged_0518 <- Reduce(
  function(x, y) left_join_drop_dups(x, y, by = by_keys),
  list(
    demo_health_0518,
    mort_0518, 
    GHG_0518,
    PHDI_0518_mean
  )
)

by_keys <- c("SEQN", "cycle", "WTDR2D")

nhanes_merged_0518 <- Reduce(
  function(x, y) left_join_drop_dups(x, y, by = by_keys),
  list(
    nhanes_merged_0518,
    HEI2020_0518,
    AHEI_0518,
    DASH_0518,
    AMED_0518,
    DII_0518
  )
)

glimpse(nhanes_merged_0518)
#View(nhanes_merged_0518)

# Save nhanes_merged_0518 to CSV file
write_csv(nhanes_merged_0518, "DATA/nhanes_merged_0518.csv")
```

```{r load nhanes_merged_0518}
# Load nhanes_merged_0518 back from CSV
nhanes_merged_0518 <- read_csv("D:/Clients/VinUni/Dr. Linh/Dietary_index/DATA/nhanes_merged_0518.csv", show_col_types = FALSE)
```

# Data cleaning

## Diet score

```{r}
# Create new dietscore variable score_cat as categorical variable (Q1, Q2, Q3, Q4) for each diet score
nhanes_merged_0518 <- nhanes_merged_0518 %>%
  mutate(
    HEI2020_cat = factor(ntile(HEI2020_ALL, 5), levels = 1:5, labels = paste0("Q", 1:5)),
    AHEI_cat    = factor(ntile(AHEI_ALL, 5), levels = 1:5, labels = paste0("Q", 1:5)),
    DASH_cat    = factor(ntile(DASH_ALL, 5), levels = 1:5, labels = paste0("Q", 1:5)),
    AMED_cat    = factor(ntile(MED_ALL, 5), levels = 1:5, labels = paste0("Q", 1:5)),
    DII_cat     = factor(ntile(DII_ALL, 5), levels = 1:5, labels = paste0("Q", 1:5)),
    PHDI_cat    = factor(ntile(PHDI_ALL_mean, 5), levels = 1:5, labels = paste0("Q", 1:5))
  )
```

## Major chronic disease (MCD) (combined variable)

### Major cardiovascular disease (M_CVD)

```{r}
### Major cardiovascular disease (M_CVD: having at least 1 disease = 1; all of these conditions are reported as “No” = 0; 
# all five conditions are missing or coded as “Refused/Don’t know”) defined as having at least one of the following conditions ###

# MCQ160C : Coronary heart disease (CHD)
# MCQ160D : Angina/angina pectoris
# MCQ160B : Congestive heart failure (CHF)
# MCQ160E : Myocardial infarction (heart attack, MI)
# MCQ160F : Stroke

# List NHANES raw variables and new labels
vars_cvd <- c("MCQ160B","MCQ160C","MCQ160D","MCQ160E","MCQ160F")
labs_cvd <- c("CHF","CHD","Angina","MI","Stroke")

# Recode each CVD variable separately using case_when: 1 = Yes, 2 = No, 7/9/. = NA
nhanes_merged_0518 <- nhanes_merged_0518 %>%
  mutate(
    # CHF (Congestive heart failure)
    CHF = case_when(
      MCQ160B == 1 ~ 1L,
      MCQ160B == 2 ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # CHD (Coronary heart disease)
    CHD = case_when(
      MCQ160C == 1 ~ 1L,
      MCQ160C == 2 ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # Angina
    Angina = case_when(
      MCQ160D == 1 ~ 1L,
      MCQ160D == 2 ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # MI (Myocardial infarction)
    MI = case_when(
      MCQ160E == 1 ~ 1L,
      MCQ160E == 2 ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # Stroke
    Stroke = case_when(
      MCQ160F == 1 ~ 1L,
      MCQ160F == 2 ~ 0L,
      TRUE ~ NA_integer_
    )
  )

# Define Major CVD using case_when
nhanes_merged_0518 <- nhanes_merged_0518 %>%
  mutate(
    M_CVD = case_when(
      # All CVD variables are NA
      is.na(CHF) & is.na(CHD) & is.na(Angina) & is.na(MI) & is.na(Stroke) ~ NA_integer_,
      # At least one CVD condition is present
      CHF == 1L | CHD == 1L | Angina == 1L | MI == 1L | Stroke == 1L ~ 1L,
      # All measured CVD conditions are 0
      TRUE ~ 0L
    )
  )
```

### Type 2 diabetes

```{r}
## Total diabetes = Diagnosed diabetes - self-reported diabetes (DIQ010 == 1) 
#                   OR Fasting plasma glucose (LBXGLU) ≥ 126 mg/dL OR HbA1c (LBXGH) ≥ 6.5%
# (AND People with missing values for either fasting glucose or A1C and pregnant women were excluded => will be included in the exclusion criteria) 

# Excluded T1D: diabetes diagnosed before the age of 30 years (DID040<30) 
#               AND who was taking only insulin therapy (DIQ050, DID060, DIQ070)
# AND during pregnancy, told you have diabetes (RHQ162==1 only in cycles from 2007) | Gestational diabetes (DIQ175S==28 only in cycles from 2011)

nhanes_merged_0518 <- nhanes_merged_0518 %>%
  mutate(
    # Diabetes: DIQ010==1 OR FPG≥126 OR HbA1c≥6.5
    diabetes = case_when(
      DIQ010 == 1 | LBXGLU >= 126 | LBXGH >= 6.5 ~ 1L,
      DIQ010 == 2 | LBXGLU < 126 | LBXGH < 6.5  ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # T1D flag: diabetes==1 & DID040<30 & ONLY insulin (DIQ050==1 AND DID060 in 1:49 AND DIQ070==2)
    T1D_flag = case_when(
      diabetes == 1L & DID040 <30 & DIQ050 == 1L & DIQ070 == 2L & DID060 %in% 1:49 ~ 1L,
      TRUE ~ 0L
    ),  
    
    # T2D: diabetes==1 AND T1D_flag==0L (NOT T1D) AND RHQ162==2 | DIQ175S!==28 (NOT gestational diabetes)
    T2D = case_when(
      diabetes == 1L & T1D_flag == 0L &  # bị ĐTĐ ko phải Type 1
        (RHQ162 == 2 | DIQ175S != 28) # ko bị tiểu đường thai kỳ  
      ~ 1L,

      diabetes == 0L # ko bị diabetes
      | RHQ162 == 1 | DIQ175S == 28 # bị tiểu đường thai kỳ 
      ~ 0L,
      
      TRUE ~ NA_integer_ 
    )
  ) #Note: If exclude pregnant women from the beginning, code should be refined
```

### Cancer

```{r}
# Recode: 1 = Yes, 2 = No, 7/9/. = NA
nhanes_merged_0518 <- nhanes_merged_0518 %>%
  mutate(
    cancer = case_when(
      MCQ220 == 1 ~ 1L,             
      MCQ220 == 2 ~ 0L,             
      TRUE ~ NA_integer_            
    )
  )
```

### MCD variable

```{r}
nhanes_merged_0518 <- nhanes_merged_0518 %>%
  mutate(
    MCD_count = rowSums(across(c(M_CVD, T2D, cancer)), na.rm = TRUE), # Count MCD
    MCD = case_when(
      rowSums(!is.na(across(c(M_CVD, T2D, cancer)))) == 0L ~ NA_integer_, # all NA
      MCD_count >= 1 ~ 1L,  
      TRUE ~ 0L             
    )
  )
```

## Metabolic syndrome (MetS)

```{r}
# defined as having at least one of the following conditions:
# 1. Abdominal obesity (Waist circumference (cm): BMXWAIST): >40 in (men), >35 in (women) => >101.6 (men), >88.9 (women)
# 2. High blood pressure (mm Hg) 
# BPXSY1:BPXSY4, BPXDI1:BPXDI4 (create average): SBP (Systolic Blood Pressure) ≥130 or DBP (Diastolic Blood Pressure) ≥80
# or BPQ020==1 (Ever told you had high blood pressure)
# 3. Impaired fasting glucose (mg/dL) (LBXGLU) : ≥100
# 4. High triglycerides (mg/dL) (LBXTR): >150 
# 5. Low HDL (HDL cholesterol (mg/dL): LBDHDD) <40 mg/dL (men), <50 mg/dL (women)
# Gender (RIAGENDR): 1 Male, 2 Female

nhanes_merged_0518 <- nhanes_merged_0518 %>%
  mutate(
    # --- Abdominal obesity (cm) ---
    wc_high = case_when(
      RIAGENDR == 1L & BMXWAIST > 101.6 ~ 1L,  # men >101.6 cm
      RIAGENDR == 2L & BMXWAIST > 88.9  ~ 1L,  # women >88.9 cm
      RIAGENDR %in% c(1L,2L) & !is.na(BMXWAIST) ~ 0L,               # measured but below threshold
      TRUE ~ NA_integer_                                            # missing waist -> NA
    ),
    
    # --- High blood pressure (use mean of 4 times of measurements) ---
    SBP_mean = {
      x <- rowMeans(as.matrix(across(any_of(c("BPXSY1","BPXSY2","BPXSY3","BPXSY4")))), na.rm = TRUE) #average available SBP readings
      ifelse(is.nan(x), NA_real_, x)   # all-NA -> NA
    },
    DBP_mean = {
      x <- rowMeans(as.matrix(across(any_of(c("BPXDI1","BPXDI2","BPXDI3","BPXDI4")))), na.rm = TRUE)
      ifelse(is.nan(x), NA_real_, x)
    },
    
    bp_high = case_when(
      SBP_mean >= 130 | DBP_mean >= 80  | BPQ020 == 1L ~ 1L,  # or self-reported history of high BP (BPQ020)
      !is.na(SBP_mean) | !is.na(DBP_mean) | BPQ020 %in% c(1L,2L) ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # --- Impaired fasting glucose ---
    glu_high = case_when(
      LBXGLU >= 100 ~ 1L,
      !is.na(LBXGLU) ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # --- High triglycerides ---
    tg_high = case_when(
      LBXTR > 150 ~ 1L,
      !is.na(LBXTR) ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # --- Low HDL ---
    hdl_low = case_when(
      RIAGENDR == 1L & LBDHDD < 40 ~ 1L,  # men
      RIAGENDR == 2L & LBDHDD < 50 ~ 1L,  # women
      RIAGENDR %in% c(1L,2L) & !is.na(LBDHDD) ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # Final Metabolic Syndrome (≥3 of 5)
    # NA if all five components are NA
    MetS_count = rowSums(across(c(wc_high, bp_high, glu_high, tg_high, hdl_low)), na.rm = TRUE),
    MetS = case_when(
      rowSums(!is.na(across(c(wc_high, bp_high, glu_high, tg_high, hdl_low)))) == 0L ~ NA_integer_,
      MetS_count >= 3 ~ 1L,
      TRUE ~ 0L
    )
  )

#View(nhanes_merged_0518)
```

## Covariates

```{r}
nhanes_merged_0518 <- nhanes_merged_0518 %>%
  mutate(
    # ---- Age group ----
    RIDAGEYR_CAT = case_when(
      RIDAGEYR >= 20 & RIDAGEYR <= 34 ~ 1L,
      RIDAGEYR >= 35 & RIDAGEYR <= 49 ~ 2L,
      RIDAGEYR >= 50 & RIDAGEYR <= 64 ~ 3L,
      RIDAGEYR >= 65                  ~ 4L,
      TRUE ~ NA_integer_
    ),
    
    # ---- Education (DMDEDUC2) ----
    DMDEDUC2_CAT = case_when(
      DMDEDUC2 %in% c(1,2) ~ 2L,   # <High School
      DMDEDUC2 == 3        ~ 3L,   # High school diploma or GED
      DMDEDUC2 == 4        ~ 4L,   # Some college or Associate’s degree
      DMDEDUC2 == 5        ~ 5L,   # College graduate or above
      DMDEDUC2 %in% c(7,9) ~ NA_integer_,
      TRUE ~ NA_integer_
    ),
    
    # ---- Marital (DMDMARTL) ----
    DMDMARTL_CAT = case_when(
      DMDMARTL == 1           ~ 1L,  # Married
      DMDMARTL %in% 2:6       ~ 0L,  # Not married (includes widowed, divorced, separated, never married, and living with a partner)
      DMDMARTL %in% c(77,99)  ~ NA_integer_,
      TRUE ~ NA_integer_
    ),
    
    # ---- Family Income to Poverty Ratio (INDFMPIR) ----
    INDFMPIR_CAT = case_when(
      INDFMPIR < 1.30   ~ 1L,
      INDFMPIR < 3.50   ~ 2L,
      INDFMPIR >= 3.50  ~ 3L,
      TRUE ~ NA_integer_
    ),
    
    # ---- BMI ----
    BMI_CAT = case_when(
      BMXBMI < 25               ~ 2L,
      BMXBMI >=25 & BMXBMI <30  ~ 3L,
      BMXBMI >=30               ~ 4L,
      TRUE ~ NA_integer_
    ),
    OBESE = case_when(
      BMXBMI >=30 ~ 1L,
      BMXBMI <30  ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # ---- Abdominal obesity ---- (already have)
    AB_OBESE = case_when(
      RIAGENDR == 1 & BMXWAIST >= 102 ~ 1L,
      RIAGENDR == 2 & BMXWAIST >=  88 ~ 1L,
      !is.na(BMXWAIST)                ~ 0L,
      TRUE ~ NA_integer_
    ),
    
    # ---- Smoking ----
    SMQ_CAT = case_when(
      SMQ020 == 2                          ~ 0L, # Never
      SMQ020 == 1 & SMQ040 == 3            ~ 1L, # Former
      SMQ020 == 1 & SMQ040 %in% c(1,2)     ~ 2L, # Current
      SMQ020 %in% c(7,9) | SMQ040 %in% c(7,9) ~ NA_integer_,
      TRUE ~ NA_integer_
    ),
    
    # ---- Alcohol ----
    ALQ130_CAT = case_when(
      ALQ130 %in% c(777,999) ~ NA_real_,
      !is.na(ALQ130)         ~ as.numeric(ALQ130),
      TRUE                   ~ NA_real_
    ),
    
    # ---- Physical activity ----
    PAQ_MINWEEK = case_when(
      !is.na(PAQ670) & !is.na(PAD675) ~ as.numeric(PAQ670) * as.numeric(PAD675),
      TRUE ~ NA_real_
    ),
    PAQ_CAT = case_when(
      PAQ_MINWEEK == 0                       ~ 0L, # Inactive
      PAQ_MINWEEK > 0 & PAQ_MINWEEK <150     ~ 1L, # Insufficient
      PAQ_MINWEEK >=150 & PAQ_MINWEEK <300   ~ 2L, # Sufficient
      PAQ_MINWEEK >=300                      ~ 3L, # Mod+vigorous
      TRUE ~ NA_integer_
    ),
    
    # ---- Supplement use ----
    DSD010_CAT = case_when(
      DSD010 == 1         ~ 1L,
      DSD010 == 2         ~ 0L,
      DSD010 %in% c(7,9)  ~ NA_integer_,
      TRUE ~ NA_integer_
    ),

    # ---- Weight ----
    WTDR14D = WTDR2D / 7
  )

### label for covariates ###
library(dplyr)

f <- function(x, lv, lab) factor(x, levels = lv, labels = lab)

nhanes_merged_0518 <- nhanes_merged_0518 %>%
  mutate(
    RIAGENDR     = f(RIAGENDR, c(1, 2), c("Male", "Female")),
    RIDAGEYR_CAT = f(RIDAGEYR_CAT, 1:4, c("20-34", "35-49", "50-64", "≥65")),
    RIDRETH1     = f(RIDRETH1, 1:5,
                     c("Mexican American","Other Hispanic","Non-Hispanic White",
                       "Non-Hispanic Black","Other Race")),
    DMDEDUC2_CAT = f(DMDEDUC2_CAT, c(2,3,4,5),
                     c("Less Than High School","High School Graduate/GED",
                       "Some College or Associate's Degree","College Graduate or Above")),
    DMDMARTL_CAT = f(DMDMARTL_CAT, c(1,0), c("Married","Not Married")),
    INDFMPIR_CAT = f(INDFMPIR_CAT, 1:3, c("0–<1.30","1.30–3.49","≥3.50")),
    BMI_CAT      = f(BMI_CAT, c(2,3,4), c("Underweight or Normal Weight","Overweight","Obese")),
    SMQ_CAT      = f(SMQ_CAT, c(0,1,2), c("Never","Former","Current")),
    PAQ_CAT      = f(PAQ_CAT, c(0,1,2,3), c("Inactive","Insufficient","Sufficient","Mod+Vigorous"))
  )
```

```{r varlabels}
var_label(nhanes_merged_0518) <- list(
  HEI2020_cat = "HEI-2020 quartile",
  AHEI_cat    = "AHEI quartile",
  DASH_cat    = "DASH quartile",
  AMED_cat    = "aMED quartile",
  DII_cat     = "DII quartile",
  PHDI_cat    = "PHDI quartile",
  M_CVD       = "Major cardiovascular disease",
  T2D         = "Type 2 diabetes",
  cancer      = "Cancer",
  MCD         = "Major chronic disease",
  MetS        = "Metabolic syndrome",
  RIAGENDR     = "Gender",
  RIDAGEYR_CAT = "Age group",
  RIDRETH1     = "Race/ethnicity",
  DMDEDUC2_CAT = "Education level",
  DMDMARTL_CAT = "Marital status",
  INDFMPIR_CAT = "Family income to poverty ratio",
  BMI_CAT      = "BMI category",
  OBESE        = "Obesity status",
  AB_OBESE     = "Abdominal obesity",
  SMQ_CAT      = "Smoking status",
  ALQ130_CAT   = "Alcohol consumption (drinks per week)",
  PAQ_CAT      = "Physical activity level",
  TOTALKCAL_PHDI_mean = "Total energy intake (kcal/day)",
  DSD010_CAT   = "Dietary supplement use",
  DMDHHSIZ    = "Household size"
)
```

## Exclusion criteria

```{r}
# participants with unreliable dietary recall as defined by NHANES: DR1DRSTZ == 1 OR DR2DRSTZ == 1 (Reliable and met the minimum criteria - Dietary recall status)
# missing day 1 or day 2 dietary recall data:  
# <=20 y old or missing age in years: RIDAGEYR >= 20 OR NA
# implausible energy intake values (for females: <600 kcal/d or >3500 kcal/d, and for males: <800 kcal/d or >4200 kcal/d): DR1TKCAL; DR2TKCAL; RIAGENDR
# DR1TKCAL; DR2TKCAL ở trên PHDI đã có ghép vào

# filter 
excluded_cols <- c("HEI2020_ALL","AHEI_ALL","DASH_ALL", "MED_ALL", "DII_ALL", "PHDI_ALL_mean")
                   #"M_CVD","T2D","cancer")
health_cols <- c("CHF","CHD","Angina","MI","Stroke",
            "M_CVD","T2D","cancer", "MCD")

nhanes_merged_0518_filtered <- nhanes_merged_0518 %>%
  filter(
    RIDAGEYR >= 20, # adults (>=20)
    if_all(all_of(excluded_cols), ~ !is.na(.)), # keep rows with NO missing in ALL excluded_cols
    if_all(all_of(health_cols), ~ !is.na(.)) # keep rows with AT LEAST ONE non-missing in excluded_cols
    # !(RIAGENDR == 2 & RIDEXPRG == 1) # exclude women pregnant
  )

# Rows: 30,883
# Rows: 27,306
```

## Missing patterns

```{r}
#| eval: false
#| include: false

# overview
# row, column
dim(nhanes_merged_0518_filtered)

# tổng số ô, tổng số missing, % missing
total_cells <- prod(dim(nhanes_merged_0518_filtered))
total_missing <- sum(is.na(nhanes_merged_0518_filtered))
missing_percentage <- total_missing / total_cells * 100

# theo biến: n_miss, pct_miss, sorted desc
miss_var <- naniar::miss_var_summary(nhanes_merged_0518_filtered) %>%
  arrange(desc(pct_miss))
miss_var %>% head(30)   # top 30 biến có missing cao

# theo case (mỗi row có bao nhiêu var missing)
miss_case <- naniar::miss_case_summary(nhanes_merged_0518_filtered) %>%
  arrange(desc(n_miss))
miss_case %>% head(20)

threshold <- 30
vars_high_missing <- miss_var %>%
  filter(pct_miss > threshold) %>%
  pull(variable)

length(vars_high_missing)
vars_high_missing 

####################################################
# % missing theo biến 
naniar::gg_miss_var(nhanes_merged_0518_filtered) +
  ggplot2::theme_minimal() +
  ggplot2::labs(title = "Percent missing per variable")

# số missing per case (hàng)
naniar::gg_miss_case(nhanes_merged_0518_filtered) +
  ggplot2::labs(title = "Missingness per observation (case)")

# ma trận missing (vis_miss) nhưng dữ liệu lớn -> sample rows
set.seed(123)
small_sample <- nhanes_merged_0518_filtered %>% slice_sample(n = 3000) # or n = 2000
naniar::vis_miss(small_sample, warn_large_data = FALSE)

# chọn top N biến có missing nhiều nhất (vd N = 40)
topN <- miss_var %>% slice_max(order_by = n_miss, n = 40) %>% pull(variable)
naniar::vis_miss(nhanes_merged_0518_filtered %>% select(all_of(topN)), warn_large_data = FALSE)



# check all variables
gg_miss_var(nhanes_merged_0518_filtered, show_pct = TRUE) + # %
  labs(y = "% missing", x = "Variables") +
  coord_flip()
gg_miss_var(nhanes_merged_0518) # n


# aggr(nhanes_merged_0518_filtered, numbers = TRUE, prop = c(TRUE, FALSE))

library(naniar)
vis_miss(nhanes_merged_0518_filtered)

miss_var_summary(nhanes_merged_0518_filtered) %>% print()
gg_miss_var(nhanes_merged_0518_filtered)

gg_miss_var(nhanes_merged_0518_filtered, show_pct = TRUE) + #%
  labs(y = "% missing", x = "Variables") +
  coord_flip()

nhanes_merged_0518_filtered %>%
  miss_var_summary() %>%
  filter(pct_miss > 0)
# missing plot removed all vars with no missing data
nhanes_merged_0518_filtered %>%
  miss_var_summary() %>%                 # -> variable, n_miss, pct_miss
  filter(pct_miss > 0) %>%               # filtered vars with no missing 
  ggplot(aes(x = pct_miss, y = reorder(variable, pct_miss))) +
  geom_segment(aes(x = 0, xend = pct_miss, yend = variable), linewidth = 0.6) +
  geom_point(size = 2) +
  labs(x = "% missing") +
  #theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank())
```

# Results

```{r list holders}
covars <- c("MCQ160B","MCQ160C","MCQ160D","MCQ160E","MCQ160F",
            "RIAGENDR","RIDAGEYR_CAT","RIDRETH1","DMDEDUC2_CAT","DMDMARTL_CAT",
            "INDFMPIR_CAT","BMI_CAT","OBESE","AB_OBESE","SMQ_CAT","PAQ_CAT",
            "M_CVD","MetS","T2D","MCD")

covars <- c("CHF","CHD","Angina","MI","Stroke",
            "M_CVD","T2D","cancer", "MCD", "MetS")

health_outcomes <- c("MCD", "M_CVD", "T2D", "cancer", "wc_high", "bp_high", "glu_high", "tg_high", "hdl_low", "MetS")
diet_scores_cont <- c("HEI2020_ALL", "AHEI_ALL", "DASH_ALL", "MED_ALL", "DII_ALL", "PHDI_ALL_mean")
diet_scores_cat <- c("HEI2020_cat", "AHEI_cat", "DASH_cat", "AMED_cat", "DII_cat", "PHDI_cat")

# Create a mapping for outcome labels
outcome_labels <- c(
  "CHF" = "Congestive Heart Failure",
  "CHD" = "Coronary Heart Disease", 
  "Angina" = "Angina/Angina Pectoris",
  "MI" = "Myocardial Infarction",
  "Stroke" = "Stroke",
  "M_CVD" = "Major Cardiovascular Disease",
  "diabetes" = "Diabetes",
  "T1D_flag" = "Type 1 Diabetes",
  "T2D" = "Type 2 Diabetes",
  "cancer" = "Cancer",
  "wc_high" = "High Waist Circumference",
  "bp_high" = "High Blood Pressure",
  "glu_high" = "High Fasting Glucose",
  "tg_high" = "High Triglycerides",
  "hdl_low" = "Low HDL Cholesterol",
  "MetS" = "Metabolic Syndrome",
  "MCD" = "Major Chronic Disease"
)
```

```{r}
nhanes_merged_0518_filter <- nhanes_merged_0518 %>%
  filter(
    !is.na(WTDR14D) & !is.na(SDMVPSU) & !is.na(SDMVSTRA)
  )
```

```{r survey design}
PHDI_GHG_FINAL_DATA_0518_design_14yr = svydesign(
    id = ~SDMVPSU, 
    strata = ~SDMVSTRA,
    weight = ~WTDR14D,
    data = nhanes_merged_0518_filter,
    nest = TRUE)

# Subpop call
design <- subset(PHDI_GHG_FINAL_DATA_0518_design_14yr, RIDAGEYR >= 20 & !is.na(get("HEI2020_cat")) & !is.na(get("AHEI_cat")) & !is.na(get("DASH_cat")) & !is.na(get("AMED_cat")) & !is.na(get("DII_cat")) & !is.na(get("PHDI_cat")) & !is.na(get("M_CVD")) & !is.na(get("T2D")) & !is.na(get("cancer")) & !is.na(get("MCD"))
)
```

```{r}
#| eval: false
#| include: false

### Describe table for health outcomes
# table for filtered data
health_des <- gt_cov_2col(nhanes_merged_0518_filtered, covars, digits = 1)
health_des
```

\newpage

```{r}
#| label: tbl-table-characteristics-adjust1
#| tbl-cap: "Adjust với gender, age, ethnic, education level, marital status,  family income to poverty ratio, smoking status, Physical activity level, Total energy intake, Alcohol consumption, Household size, Dietary supplement use"

# Define outcomes and diet scores
outcomes <- c("MCD", "M_CVD", "T2D", "cancer", "wc_high", "bp_high", "glu_high", "tg_high", "hdl_low", "MetS")
diet_scores_cont <- c("HEI2020_ALL", "AHEI_ALL", "DASH_ALL", "MED_ALL", "DII_ALL", "PHDI_ALL_mean")
diet_scores_cat <- c("HEI2020_cat", "AHEI_cat", "DASH_cat", "AMED_cat", "DII_cat", "PHDI_cat")
adjust <- " + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDEDUC2_CAT + DMDMARTL_CAT + INDFMPIR_CAT + SMQ_CAT + PAQ_CAT + TOTALKCAL_PHDI_mean + ALQ130_CAT + DMDHHSIZ + DSD010"

# Function to create table for one outcome and one diet score
create_outcome_table <- function(outcome, ds_cont, ds_cat, design_obj) {
  
  # Initialize results data frame
  results <- data.frame(
    Quintile = c(paste0("Q", 1:5), "Total"),
    Median_IQR = NA_character_,
    N_Cases = NA_character_,
    stringsAsFactors = FALSE
  )
  
  # Get underlying data for unweighted counts
  data_df <- design_obj$variables
  
  # Calculate median/IQR and case counts for each quintile
  for (q in 1:5) {
    quintile_label <- paste0("Q", q)
    
    # Subset design for this quintile AND outcome cases
    design_q <- subset(design_obj, get(ds_cat) == quintile_label & get(outcome) == 1)
    
    # 1. Median and IQR of continuous diet score among cases
    median_val <- svyquantile(as.formula(paste0("~", ds_cont)), design_q, quantiles = 0.5)[[1]]
    q25 <- svyquantile(as.formula(paste0("~", ds_cont)), design_q, quantiles = 0.25)[[1]]
    q75 <- svyquantile(as.formula(paste0("~", ds_cont)), design_q, quantiles = 0.75)[[1]]
    results$Median_IQR[q] <- sprintf("%.1f (%.1f-%.1f)", median_val, q25, q75)
    
    # 2. Number of cases (unweighted)
    n_cases_unweighted <- sum(data_df[[ds_cat]] == quintile_label & data_df[[outcome]] == 1, na.rm = TRUE)
    results$N_Cases[q] <- sprintf("%d", n_cases_unweighted)
  }
  
  # Total row calculations
  design_total <- subset(design_obj, get(outcome) == 1)
  median_total <- svyquantile(as.formula(paste0("~", ds_cont)), design_total, quantiles = 0.5)[[1]]
  q25_total <- svyquantile(as.formula(paste0("~", ds_cont)), design_total, quantiles = 0.25)[[1]]
  q75_total <- svyquantile(as.formula(paste0("~", ds_cont)), design_total, quantiles = 0.75)[[1]]
  results$Median_IQR[6] <- sprintf("%.1f (%.1f-%.1f)", median_total, q25_total, q75_total)
  n_cases_total <- sum(data_df[[outcome]] == 1, na.rm = TRUE)
  results$N_Cases[6] <- sprintf("%d", n_cases_total)
  
  # Run age-adjusted model once for all quintiles
  formula_age <- as.formula(paste0(outcome, " ~ ", ds_cat, " + RIDAGEYR"))
  model_age <- svyglm(formula_age, design = design_obj, family = quasibinomial())
  coef_age <- summary(model_age)$coefficients
  n_obs_age <- model_age$df.null + 1  # Number of observations in age-adjusted model
  
  # Extract age-adjusted ORs for each quintile
  results$OR_Age <- NA_character_
  results$OR_Age[1] <- "1.00 (ref)"
  results$OR_Age[6] <- sprintf("N=%d", n_obs_age)  # Show sample size in total row
  for (q in 2:5) {
    row_name <- paste0(ds_cat, "Q", q)
    if (row_name %in% rownames(coef_age)) {
      or_age <- exp(coef_age[row_name, "Estimate"])
      se_age <- coef_age[row_name, "Std. Error"]
      ci_low_age <- exp(coef_age[row_name, "Estimate"] - 1.96 * se_age)
      ci_high_age <- exp(coef_age[row_name, "Estimate"] + 1.96 * se_age)
      results$OR_Age[q] <- sprintf("%.2f (%.2f-%.2f)", or_age, ci_low_age, ci_high_age)
    }
  }
  
  # Run fully adjusted model once for all quintiles
  formula_full <- as.formula(paste0(
    outcome, " ~ ", ds_cat, adjust
  ))
  model_full <- svyglm(formula_full, design = design_obj, family = quasibinomial())
  coef_full <- summary(model_full)$coefficients
  n_obs_full <- model_full$df.null + 1  # Number of observations in fully adjusted model
  
  # Extract fully adjusted ORs for each quintile
  results$OR_Full <- NA_character_
  results$OR_Full[1] <- "1.00 (ref)"
  results$OR_Full[6] <- sprintf("N=%d", n_obs_full)  # Show sample size in total row
  for (q in 2:5) {
    row_name <- paste0(ds_cat, "Q", q)
    if (row_name %in% rownames(coef_full)) {
      or_full <- exp(coef_full[row_name, "Estimate"])
      se_full <- coef_full[row_name, "Std. Error"]
      ci_low_full <- exp(coef_full[row_name, "Estimate"] - 1.96 * se_full)
      ci_high_full <- exp(coef_full[row_name, "Estimate"] + 1.96 * se_full)
      results$OR_Full[q] <- sprintf("%.2f (%.2f-%.2f)", or_full, ci_low_full, ci_high_full)
    }
  }
  
  return(results)
}

# Create table for each outcome and diet score combination
all_tables <- list()

for (outcome in outcomes) {
  for (i in 1:length(diet_scores_cont)) {
    ds_cont <- diet_scores_cont[i]
    ds_cat <- diet_scores_cat[i]
    
    cat("\nProcessing:", outcome, "-", ds_cont, "\n")
    
    tryCatch({
      table_result <- create_outcome_table(outcome, ds_cont, ds_cat, design)
      
      # Add outcome and diet score labels
      table_result <- cbind(
        Outcome = outcome_labels[outcome],
        Diet_Score = gsub("_ALL|_mean", "", ds_cont),
        table_result
      )
      
      all_tables[[paste(outcome, ds_cat, sep = "_")]] <- table_result
    }, error = function(e) {
      cat("Error:", e$message, "\n")
    })
  }
}

# Combine all results
final_table <- do.call(rbind, all_tables)
rownames(final_table) <- NULL

# Create flextable
ft <- flextable(final_table) %>%
  set_header_labels(
    Outcome = "Outcome",
    Diet_Score = "Diet Score",
    Quintile = "Quintile",
    Median_IQR = "Median (IQR)",
    N_Cases = "Cases (n)",
    OR_Age = "Age-adjusted OR (95% CI)",
    OR_Full = "Fully adjusted OR (95% CI)"
  ) %>%
  merge_v(j = c("Outcome", "Diet_Score")) %>%
  theme_box() %>%
  autofit() %>%
  FitFlextableToPage(pgwidth = 6.5)

ft
```

\newpage

```{r}
#| label: tbl-table-characteristics-adjust2
#| tbl-cap: "Adjust với gender, age, ethnic, education level, marital status,  family income to poverty ratio, smoking status, Physical activity level, Total energy intake, Alcohol consumption, Household size, Dietary supplement use"

# Define outcomes and diet scores
outcomes <- c("MCD", "M_CVD", "T2D", "cancer", "wc_high", "bp_high", "glu_high", "tg_high", "hdl_low", "MetS")
diet_scores_cont <- c("HEI2020_ALL", "AHEI_ALL", "DASH_ALL", "MED_ALL", "DII_ALL", "PHDI_ALL_mean")
diet_scores_cat <- c("HEI2020_cat", "AHEI_cat", "DASH_cat", "AMED_cat", "DII_cat", "PHDI_cat")
adjust <- " + RIAGENDR + RIDAGEYR + RIDRETH1 + DMDEDUC2_CAT + DMDMARTL_CAT + SMQ_CAT + TOTALKCAL_PHDI_mean + DMDHHSIZ"

# Function to create table for one outcome and one diet score
create_outcome_table <- function(outcome, ds_cont, ds_cat, design_obj) {
  
  # Initialize results data frame
  results <- data.frame(
    Quintile = c(paste0("Q", 1:5), "Total"),
    Median_IQR = NA_character_,
    N_Cases = NA_character_,
    stringsAsFactors = FALSE
  )
  
  # Get underlying data for unweighted counts
  data_df <- design_obj$variables
  
  # Calculate median/IQR and case counts for each quintile
  for (q in 1:5) {
    quintile_label <- paste0("Q", q)
    
    # Subset design for this quintile AND outcome cases
    design_q <- subset(design_obj, get(ds_cat) == quintile_label & get(outcome) == 1)
    
    # 1. Median and IQR of continuous diet score among cases
    median_val <- svyquantile(as.formula(paste0("~", ds_cont)), design_q, quantiles = 0.5)[[1]]
    q25 <- svyquantile(as.formula(paste0("~", ds_cont)), design_q, quantiles = 0.25)[[1]]
    q75 <- svyquantile(as.formula(paste0("~", ds_cont)), design_q, quantiles = 0.75)[[1]]
    results$Median_IQR[q] <- sprintf("%.1f (%.1f-%.1f)", median_val, q25, q75)
    
    # 2. Number of cases (unweighted)
    n_cases_unweighted <- sum(data_df[[ds_cat]] == quintile_label & data_df[[outcome]] == 1, na.rm = TRUE)
    results$N_Cases[q] <- sprintf("%d", n_cases_unweighted)
  }
  
  # Total row calculations
  design_total <- subset(design_obj, get(outcome) == 1)
  median_total <- svyquantile(as.formula(paste0("~", ds_cont)), design_total, quantiles = 0.5)[[1]]
  q25_total <- svyquantile(as.formula(paste0("~", ds_cont)), design_total, quantiles = 0.25)[[1]]
  q75_total <- svyquantile(as.formula(paste0("~", ds_cont)), design_total, quantiles = 0.75)[[1]]
  results$Median_IQR[6] <- sprintf("%.1f (%.1f-%.1f)", median_total, q25_total, q75_total)
  n_cases_total <- sum(data_df[[outcome]] == 1, na.rm = TRUE)
  results$N_Cases[6] <- sprintf("%d", n_cases_total)
  
  # Run age-adjusted model once for all quintiles
  formula_age <- as.formula(paste0(outcome, " ~ ", ds_cat, " + RIDAGEYR"))
  model_age <- svyglm(formula_age, design = design_obj, family = quasibinomial())
  coef_age <- summary(model_age)$coefficients
  n_obs_age <- model_age$df.null + 1  # Number of observations in age-adjusted model
  
  # Extract age-adjusted ORs for each quintile
  results$OR_Age <- NA_character_
  results$OR_Age[1] <- "1.00 (ref)"
  results$OR_Age[6] <- sprintf("N=%d", n_obs_age)  # Show sample size in total row
  for (q in 2:5) {
    row_name <- paste0(ds_cat, "Q", q)
    if (row_name %in% rownames(coef_age)) {
      or_age <- exp(coef_age[row_name, "Estimate"])
      se_age <- coef_age[row_name, "Std. Error"]
      ci_low_age <- exp(coef_age[row_name, "Estimate"] - 1.96 * se_age)
      ci_high_age <- exp(coef_age[row_name, "Estimate"] + 1.96 * se_age)
      results$OR_Age[q] <- sprintf("%.2f (%.2f-%.2f)", or_age, ci_low_age, ci_high_age)
    }
  }
  
  # Run fully adjusted model once for all quintiles
  formula_full <- as.formula(paste0(
    outcome, " ~ ", ds_cat, adjust
  ))
  model_full <- svyglm(formula_full, design = design_obj, family = quasibinomial())
  coef_full <- summary(model_full)$coefficients
  n_obs_full <- model_full$df.null + 1  # Number of observations in fully adjusted model
  
  # Extract fully adjusted ORs for each quintile
  results$OR_Full <- NA_character_
  results$OR_Full[1] <- "1.00 (ref)"
  results$OR_Full[6] <- sprintf("N=%d", n_obs_full)  # Show sample size in total row
  for (q in 2:5) {
    row_name <- paste0(ds_cat, "Q", q)
    if (row_name %in% rownames(coef_full)) {
      or_full <- exp(coef_full[row_name, "Estimate"])
      se_full <- coef_full[row_name, "Std. Error"]
      ci_low_full <- exp(coef_full[row_name, "Estimate"] - 1.96 * se_full)
      ci_high_full <- exp(coef_full[row_name, "Estimate"] + 1.96 * se_full)
      results$OR_Full[q] <- sprintf("%.2f (%.2f-%.2f)", or_full, ci_low_full, ci_high_full)
    }
  }
  
  return(results)
}

# Create table for each outcome and diet score combination
all_tables <- list()

for (outcome in outcomes) {
  for (i in 1:length(diet_scores_cont)) {
    ds_cont <- diet_scores_cont[i]
    ds_cat <- diet_scores_cat[i]
    
    cat("\nProcessing:", outcome, "-", ds_cont, "\n")
    
    tryCatch({
      table_result <- create_outcome_table(outcome, ds_cont, ds_cat, design)
      
      # Add outcome and diet score labels
      table_result <- cbind(
        Outcome = outcome_labels[outcome],
        Diet_Score = gsub("_ALL|_mean", "", ds_cont),
        table_result
      )
      
      all_tables[[paste(outcome, ds_cat, sep = "_")]] <- table_result
    }, error = function(e) {
      cat("Error:", e$message, "\n")
    })
  }
}

# Combine all results
final_table <- do.call(rbind, all_tables)
rownames(final_table) <- NULL

# Create flextable
ft <- flextable(final_table) %>%
  set_header_labels(
    Outcome = "Outcome",
    Diet_Score = "Diet Score",
    Quintile = "Quintile",
    Median_IQR = "Median (IQR)",
    N_Cases = "Cases (n)",
    OR_Age = "Age-adjusted OR (95% CI)",
    OR_Full = "Fully adjusted OR (95% CI)"
  ) %>%
  merge_v(j = c("Outcome", "Diet_Score")) %>%
  theme_box() %>%
  autofit() %>%
  FitFlextableToPage(pgwidth = 6.5)

ft
```

\newpage

```{r}
#| eval: false
#| include: false
#| label: tbl-characteristics
#| tbl-cap: "Characteristics of the study population"

tables_list <- list()
vars_to_include <- c(
  "RIAGENDR","RIDAGEYR_CAT","RIDRETH1","DMDEDUC2_CAT","DMDMARTL_CAT",
  "INDFMPIR_CAT","SMQ_CAT","PAQ_CAT","TOTALKCAL_PHDI_mean","ALQ130_CAT",
  "DMDHHSIZ","DSD010_CAT"
)

for (ds in diet_scores_cat) {
  # ensure ordered two-level factor for by
  by_sym <- rlang::sym(ds)

  tables_list[[ds]] <-
    gtsummary::tbl_svysummary(
      design,
      by = !!by_sym,
      include = tidyselect::all_of(vars_to_include),
      type = list(DMDHHSIZ ~ "continuous"),
      statistic = list(all_continuous() ~ "{mean} ({sd})")
    ) |>
    add_p() |>
    add_overall() |>
    as_flex_table() |>
    FitFlextableToPage(pgwidth = 6)
}

# Display all in one table
tables_list$HEI2020_cat
tables_list$AHEI_cat
tables_list$DASH_cat
tables_list$AMED_cat
tables_list$DII_cat
tables_list$PHDI_cat
```

\newpage

```{r}
#| eval: false
#| include: false
#| label: tbl-table-multivariate-dietscore-ncds
#| tbl-cap: "Multivariable logistic regression between diet scores and major chronic diseases (Q5 vs Q1)"

out_list_multivariable <- list()

for (outcome in health_outcomes) {
  for (ds in diet_scores_cat) {
    # Fully adjusted model with survey design
    formula <- as.formula(paste0(
      outcome, " ~ ", ds, adjust
    ))
    formula_age <- as.formula(paste0(outcome, " ~ ", ds, " + RIDAGEYR"))

    model <- svyglm(formula_age, design = design, family = quasibinomial())
    
    tidy_model <- broom::tidy(model) %>%
      mutate(
        Outcome = outcome_labels[outcome],
        Diet_Score = gsub("_cat", "", ds),
        OR = exp(estimate),
        Lower_CI = exp(estimate - 1.96 * std.error),
        Upper_CI = exp(estimate + 1.96 * std.error)
      ) %>%
      select(Outcome, Diet_Score, term, p.value, OR, Lower_CI, Upper_CI) %>%
      filter(grepl("Q5", term))  # Only keep Q5 coefficient
    
    # Use unique key combining outcome and diet score
    out_list_multivariable[[paste(outcome, ds, sep = "_")]] <- tidy_model
  }
}
results_df_multivariable <- bind_rows(out_list_multivariable)

# Create flextable
results_flextable_multivariable <- results_df_multivariable %>%
  mutate(
    OR = ifelse(OR > 10000, NA_real_, OR),
    Lower_CI = ifelse(Lower_CI > 10000, NA_real_, Lower_CI),
    Upper_CI = ifelse(Upper_CI > 10000, NA_real_, Upper_CI)
  ) %>%
  flextable() %>%
  set_header_labels(
    Outcome = "Health Outcome",
    Diet_Score = "Diet Score",
    term = "Comparison",
    OR = "Odds Ratio (OR)",
    Lower_CI = "95% CI Lower",
    Upper_CI = "95% CI Upper",
    p.value = "p-value"
  ) %>%
  colformat_double(j = c("OR", "Lower_CI", "Upper_CI"), digits = 2) %>%
  colformat_double(j = "p.value", digits = 4) %>%
  FitFlextableToPage(pgwidth = 6.5)

results_flextable_multivariable
```

\newpage

```{r}
#| eval: false
#| include: false
#| label: fig-forest
#| fig-cap: "Forest plot multivariable-adjusted logistic regression of dietary pattern scores with risk of major chronic disease"
#| fig-width: 16
#| fig-height: 12

# Prepare forest plot data - filter to only the diet score variables (tertile 5 vs 1)
forest_data <- results_df_multivariable %>%
  filter(grepl("_catQ5", term, ignore.case = TRUE)) %>%
  mutate(
    # Create cleaner diet score labels
    Diet_Score_Label = case_when(
      grepl("HEI2020", Diet_Score) ~ "HEI2020",
      grepl("AHEI", Diet_Score) ~ "AHEI",
      grepl("DASH", Diet_Score) ~ "DASH",
      grepl("AMED", Diet_Score) ~ "AMED",
      grepl("DII", Diet_Score) ~ "DII",
      grepl("PHDI", Diet_Score) ~ "PHDI",
      TRUE ~ Diet_Score
    ),

    # Calculate the EValue
    EValue = case_when(
      OR >= 1 & (Outcome == "Major Cardiovascular Disease" | Outcome == "Type 2 Diabetes" | Outcome == "Cancer") ~ OR + sqrt(OR * (OR - 1)),
      OR < 1 & (Outcome == "Major Cardiovascular Disease" | Outcome == "Type 2 Diabetes" | Outcome == "Cancer") ~ (1/OR) + sqrt((1/OR) * ((1/OR) - 1)),
      Outcome == "Major Chronic Disease" ~ sqrt(OR)
    ),

    # Clean up diet score labels to match the example
    Diet_Score_Label = factor(Diet_Score_Label, 
                               levels = c("HEI2020", "AHEI", "DASH", "AMED", "DII", "PHDI")),
    # Format OR and CI for display
    OR_CI_text = sprintf("%.2f (%.2f-%.2f)", OR, Lower_CI, Upper_CI)
  ) %>%
  arrange(Outcome, Diet_Score_Label)

# Create separate plots for each outcome
plot_list <- list()

for (outcome_name in unique(forest_data$Outcome)) {
  outcome_data <- forest_data %>% filter(Outcome == outcome_name)
  
  # Calculate y positions for text (one for each diet score)
  n_scores <- nrow(outcome_data)
  y_positions <- seq(n_scores, 1, by = -1)
  
  # Create the plot with extended negative x-axis for text
  p <- ggplot(outcome_data, aes(y = fct_rev(Diet_Score_Label), x = OR, color = Diet_Score_Label)) +
    # Add vertical reference line at OR = 1
    geom_vline(xintercept = 1, linetype = "dashed", color = "gray40", linewidth = 0.5) +
    # Error bars for 95% CI
    geom_errorbarh(aes(xmin = Lower_CI, xmax = Upper_CI, color = Diet_Score_Label), 
                   height = 0.2, linewidth = 0.8) +
    # Points for OR
    geom_point(size = 3) +
    # Add OR (95% CI) text in left margin (negative x values)
    geom_text(aes(label = OR_CI_text), 
              x = -1.6, hjust = 0, size = 3.5, color = "black", fontface = "plain") +
    # Add E-value text in left margin (further left)
    geom_text(aes(label = sprintf("%.2f", EValue)), 
              x = -0.3, hjust = 0, size = 3.5, color = "black", fontface = "plain") +
    # X-axis scale - extended to negative for text space
    scale_x_continuous(
      limits = c(-1.8, 2.8),
      breaks = c(0.5, 1.0, 1.5, 2.0, 2.5),
      labels = c("0.5", "1.0", "1.5", "2.0", "2.5"),
      expand = c(0, 0),
    ) +
    # Colors for each diet score
    scale_color_manual(values = c(
      "HEI2020" = "#E74C3C",
      "AHEI" = "#F39C12", 
      "DASH" = "#27AE60",
      "AMED" = "#95A5A6",
      "DII" = "#3498DB",
      "PHDI" = "#9B59B6"
    )) +
    # Labels
    labs(
      title = outcome_name,
      x = "OR",
      y = NULL
    ) +
    # Add column headers in left margin
    annotate("text", x = -1.6, y = n_scores + 0.8, 
             label = "OR (95% CI)", 
             hjust = 0, size = 3.5, fontface = "bold") +
    annotate("text", x = -0.4, y = n_scores + 0.8, 
             label = "E-value", 
             hjust = 0, size = 3.5, fontface = "bold") +
    # Theme
    theme_minimal() +
    theme(
      legend.position = "none",
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line.x = element_line(color = "black", linewidth = 0.5),
      axis.ticks.x = element_line(color = "black", linewidth = 0.5),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
      axis.title.x = element_text(size = 10, face = "bold"),
      axis.text.y = element_text(size = 10, face = "bold"),
      axis.text.x = element_text(size = 9),
      plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
    )
  
  plot_list[[outcome_name]] <- p
}

# Arrange plots in a grid
gridExtra::grid.arrange(grobs = plot_list, ncol = 2)
```

\newpage

```{r}
#| eval: false
#| include: false
#| label: fig-splines
#| fig-cap: "Multivariable-adjusted spline analysis of dietary pattern scores with risk of major chronic disease"
#| fig-width: 16
#| fig-height: 12

# Define diet scores and outcome
outcome <- "mortstat"
diet_scores_cont <- c("HEI2020_ALL", "AHEI_ALL", "DASH_ALL", "MED_ALL", "DII_ALL", "PHDI_ALL_mean")

# Define nice labels for plotting
score_labels <- c(
  "HEI2020_ALL" = "HEI-2020",
  "AHEI_ALL" = "AHEI", 
  "DASH_ALL" = "DASH",
  "MED_ALL" = "Mediterranean",
  "DII_ALL" = "DII",
  "PHDI_ALL_mean" = "PHDI"
)

# Initialize list to store all plot data
all_plot_data <- list()

# Loop through each dietary score
for (ds in diet_scores_cont) {
  
  # Create formula with current dietary score (continuous, no splines)
  formula_cox <- as.formula(paste("Surv(RIDAGEYR, ", "mortstat > 0", ") ~", ds, " + RIAGENDR + RIDRETH1 + DMDEDUC2_CAT + DMDMARTL_CAT + INDFMPIR_CAT + SMQ_CAT + TOTALKCAL_PHDI_mean + ALQ130_CAT + DMDHHSIZ + DSD010"))
  
  # Fit the Cox model
  model_cox <- svycoxph(formula_cox, design = design)
  
  # Extract coefficient and confidence interval for the dietary score
  coef_summary <- summary(model_cox)
  diet_coef_row <- which(rownames(coef_summary$coefficients) == ds)
  
  if (length(diet_coef_row) > 0) {
    # Get coefficient, HR, and confidence intervals
    coef_estimate <- coef_summary$coefficients[diet_coef_row, "coef"]
    hr_estimate <- exp(coef_estimate)
    hr_lower <- coef_summary$conf.int[diet_coef_row, "lower .95"]
    hr_upper <- coef_summary$conf.int[diet_coef_row, "upper .95"]
    
    # Generate score range
    score_range <- seq(min(nhanes_merged_0518_filtered[[ds]], na.rm = TRUE),
                      max(nhanes_merged_0518_filtered[[ds]], na.rm = TRUE), 
                      length.out = 100)
    
    # Calculate HR for each score value (assuming linear relationship)
    # HR = exp(coefficient * (score - reference_score))
    # Using minimum score as reference (HR = 1 at minimum)
    reference_score <- min(nhanes_merged_0518_filtered[[ds]], na.rm = TRUE)
    
    plot_data <- data.frame(
      score_value = score_range,
      hr = exp(coef_estimate * (score_range - reference_score)),
      hr_lower = exp((coef_estimate - 1.96 * coef_summary$coefficients[diet_coef_row, "se(coef)"]) * 
                    (score_range - reference_score)),
      hr_upper = exp((coef_estimate + 1.96 * coef_summary$coefficients[diet_coef_row, "se(coef)"]) * 
                    (score_range - reference_score)),
      diet_score = score_labels[ds]
    )
    
    all_plot_data[[ds]] <- plot_data
  }
}

# Combine all data
combined_data <- bind_rows(all_plot_data)

# Create the combined plot
ggplot(combined_data, aes(x = score_value, y = hr, color = diet_score)) +
  geom_line(linewidth = 1.2) +
  geom_ribbon(aes(ymin = hr_lower, ymax = hr_upper, fill = diet_score), alpha = 0.2, color = NA) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "black", alpha = 0.7) +
  facet_wrap(~ diet_score, scales = "free_x", ncol = 3) +
  labs(
    x = "Dietary Score",
    y = "Hazard Ratio (HR)"
  ) +
  scale_color_brewer(type = "qual", palette = "Set2", guide = "none") +
  scale_fill_brewer(type = "qual", palette = "Set2", guide = "none") +
  scale_y_continuous(
    limits = c(0, 2),
    breaks = seq(0.5, 2, by = 0.25),
    expand = c(0.02, 0)
  ) +
  scale_x_continuous(expand = c(0.02, 0)) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 9),
    strip.text = element_text(size = 11, face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.5)
  )
```

\newpage

```{r}
#| eval: false
#| include: false

# Test each covariate individually
test_vars <- c("RIAGENDR", "RIDAGEYR_CAT", "RIDRETH1", "DMDEDUC2_CAT", "DMDMARTL_CAT", "INDFMPIR_CAT", "SMQ_CAT", "PAQ_CAT", "TOTALKCAL_PHDI_mean", "ALQ130", "DSD010_CAT", "DMDHHSIZ")

for(var in test_vars) {
  tryCatch({
    formula_test <- as.formula(paste("CHF ~", var))
    model_test <- glm(formula_test, data = nhanes_merged_0518_filtered, family = binomial)
    cat("✓", var, "is OK\n")
  }, error = function(e) {
    cat("✗", var, "has an issue:", e$message, "\n")
  })
}
```